\begin{vcenterpage}
{\LARGE{\sc Acknowledgements}}

\noindent\rule[2pt]{\textwidth}{0.5pt}

Big thanks go to my advisor Dr. Ram\'on Carriles for his exceptional patience and guidance throughout my extremely lethargic writing process. It is possible that his patience knows no bounds. His contribution to this work can be appreciated on every page.

I extend a huge acknowledgement to Dr. Mike Downer at UT Austin for allowing me to be a part of his research group for a month, and for the extra financial support he provided for me at no benefit to him. 

Very special thanks go to Junwei Wei, Ph.D. student at UT Austin for giving up his time helping me with the experimental part of my work. Without him none of this would have been possible. His skill and ability in the lab are amazing.

Dr. Bernardo Mendoza and Dr. Enrique Castro deserve great merit for accepting the torturous job of reviewing my work. I'd also like to thank Bernardo for accepting me into his Ph.D. group even after this!

My thanks go to Dr. Alejandro Reyes Esqueda of the UNAM for providing the samples used in this research.

I am very grateful towards the CONACyT for the financial support that allows a foreigner like me to study in an excellent institution like the CIO. I feel more at home in Mexico thanks to both institutions.

I would like to thank my close friends and blood brothers for all the good times had while I was arduously working on this thesis. Juan Jes\'us S\'anchez, Sergio Romero, Jos\'e Alberto Aguilar, and Marcelo Pereira kept me alert and ever vigilant -- whether in the office or at the \emph{Chemita}.

I also wish to thank my parents, Mike and Ana, for the considerable support they have given me throughout the years. They have always supported my endeavors in all possible ways.

Lastly, my most heartfelt gratitude goes towards my girlfriend Edith who motivated me to continue down this path in the first place. She has always given me the support and love needed to complete any goal I have set for myself, and I know that I can overcome any obstacle with her.

\noindent\rule[2pt]{\textwidth}{0.5pt}
\end{vcenterpage}
